\section{Timing Compartments}
    Timing compartments are a new architecture primitive that control timing 
    channel leakage among software entities that share hardware. When combined 
    with explicit channel protection (such as access controls) timing 
    compartments provide total software isolation.
    A timing compartment consists of one or more software entities. Here a 
    software entitiy is some system abstraction (such as processes or threads 
    in a single OS system or virtual machines in a virtualization based system) 
    that execute software and have an owner.
    
    Timing channels between timing compartments must be controlled according to 
    a policy that meets the security requirements of the system. This policy 
    entails of a lattice of security levels such as $\mathtt{low} \leq
    \mathtt{high}$. A timing compartment can only leak information to a 
    preceeding timing compartment in the lattice. This lattice model is quite 
    expressive. The lattice $\mathtt{low} \leq \mathtt{high}$ defines a policy 
    where a high assurance software entity can not leak information to a low 
    assurance one. The lattice $\mathtt{T_1} \nleq \mathtt{T_2}, \mathtt{T_2} 
    \nleq \mathtt{T_1}$ defines a policy where $T_1$ and $T_2$ are mutually 
    distrusting. By convention, security lables share names with their timing 
    compartments. 
 
    To enforce the policy a trusted software component called the timing 
    compartment manager (TCM) confines software entities into TCs. The TCM then 
    informs the hardware of the TCs and policy. At runtime, the TCM tags 
    hardware requests with a tag to indicate the TC of the software entity that 
    made the request. 

    Timing compartments only address timing channels; they do not control 
    information flow through explicit channels. Handling these concerns 
    separately allows for more flexibility in the overall system design.  When 
    designing a secure system, implementors must consider how the cost required 
    to carry out a particular attack compares with other attacks, the potential 
    damage that could be caused by an attack, and the cost and performance 
    impact of implementing the security mechanisms needed to stop it. This 
    enables timing compartments to provide timing channel protection only to 
    software entities that need it.

